\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-16 manuscript/main.tex
scalar multiple argument

	\subsection{Obtaining the Estimated Hybrid Regression Coefficient}
	\label{subsec: Obtaining the Estimated Hybrid Regression Coefficient}
	The regression coefficient $\boldsymbol{\eta}$ in model \eqref{eq: Hybrid functional model} can be written in terms of the estimated PLS components and scores. First, note that we can show $\widehat{\rho}_{il} = \langle \mathbf{W}_i^{[l]}, \widehat{\xi}_l \rangle_\mathcal{H} = \langle \mathbf{W}_i, \widehat{\boldsymbol{\xi}_{K+1}}_l \rangle_\mathcal{H}$, where $\widehat{\boldsymbol{\xi}_{K+1}}_l = \widehat{\xi}_l - \sum_{u=1}^{l-1} \langle \widehat{\boldsymbol{\delta}}_u, \widehat{\xi}_l \rangle_{\mathcal{H}} \widehat{\boldsymbol{\xi}_{K+1}}_u$ for $l \ge 2$, with $\widehat{\boldsymbol{\xi}_{K+1}}_1 = \widehat{\xi}_1$. Then, the decomposition of $Y_i$ in \eqref{eq: decomposition} leads to
	\begin{equation*}
		Y_i = \sum \limits_{l=1}^L \widehat{\nu}_l \langle \mathbf{W}_i^{[l]}, \widehat{\xi}_l \rangle_\mathcal{H} +\epsilon_i = \left\langle \mathbf{W}_i, \sum \limits_{l=1}^L\widehat{\nu}_l \widehat{\boldsymbol{\xi}_{K+1}}_l \right\rangle_\mathcal{H}+\epsilon_i,
	\end{equation*}
	which, given the uniqueness of $\boldsymbol{\eta}$, leads to
	\begin{equation*}
		\widehat{\boldsymbol{\eta}} = \sum \limits_{l=1}^L\widehat{\nu}_l \widehat{\boldsymbol{\xi}_{K+1}}_l.
	\end{equation*}
	Here, the number of PLS components $L$ to be estimated can be chosen by cross-validation.

@@

	Let $\xi = (\xi_1, \ldots, \xi_K, \boldsymbol{\xi}_{K+1}) \in \mathcal{H}$. Assume that each functional component $\xi_j \in L^2([0,1])$ lies in the span of the basis functions ${b_1(t), \ldots, b_M(t)}$ and can be written as

@@

		\boldsymbol{\xi}_{K+1} 

@@

		( \mathbf{Z}_i^\top \boldsymbol{\xi}_{K+1} )

@@

		Z \boldsymbol{\xi}_{K+1} 

@@

		\boldsymbol{\xi}_{K+1}^\top \boldsymbol{\xi}_{K+1}

@@

		\boldsymbol{\xi}_{K+1}^\top \boldsymbol{\xi}_{K+1},

@@

 \section{Proof of Proposition  \ref{proposition:linear_noregul}}\label{section:proof:proposition:linear_noregul}
 \begin{proof}
 	The optimization problem
 	\ref{eq: Regularized generalized rayleigh quotient equation} can be written as
 		\begin{align*}
 		\max_{\boldsymbol{\gamma}_1, \ldots, \boldsymbol{\gamma}_K \in \mathbb{R}^M,
 			\boldsymbol{\xi}_{K+1} \in \mathbb{R}^p
 		}~&
 		\frac{1}{n^2}
 		\biggl(
 		\sum \limits_{k=1}^K 
 		\mathbf{y}^\top
 		\Theta_{j}
 		B  
 		\boldsymbol{\gamma}_j
 		+
 		\mathbf{y}^\top
 		Z \boldsymbol{\xi}_{K+1} 
 		\biggr)^2,
 		\\\text{subject to}\quad&	\sum \limits_{j=1}^K 
 		\boldsymbol{\gamma}_j^\top
 	(	B + \lambda_k B^{\prime \prime})
 		\boldsymbol{\gamma}_j
 		+
 		\boldsymbol{\xi}_{K+1}^\top \boldsymbol{\xi}_{K+1} = 1.
 	\end{align*} 
 Let $\mathbf{u}_j := \Theta_j^\top \mathbf{y} \in \mathbb{R}^M$ and $\mathbf{v} := Z^\top \mathbf{y} \in \mathbb{R}^p$. These are fixed problem data and not decision variables. Then the optimization problem becomes
 	 		\begin{align*}
 		\max_{\boldsymbol{\gamma}_1, \ldots, \boldsymbol{\gamma}_K \in \mathbb{R}^M,
 			\boldsymbol{\xi}_{K+1} \in \mathbb{R}^p
 		}~&
 		\frac{1}{n^2}
 		\biggl(
 		\sum \limits_{k=1}^K 
 	\mathbf{u}_j^\top
 		B  
 		\boldsymbol{\gamma}_j
 		+
 		\mathbf{v}^\top \boldsymbol{\xi}_{K+1} 
 		\biggr)^2,
 		\\\text{subject to}\quad&	\sum \limits_{j=1}^K 
 		\boldsymbol{\gamma}_j^\top
 		(	B + \lambda_k B^{\prime \prime})
 		\boldsymbol{\gamma}_j
 		+
 		\boldsymbol{\xi}_{K+1}^\top \boldsymbol{\xi}_{K+1} = 1.
 	\end{align*}  	
The feasible set is compact, as it corresponds to the boundary of an ellipsoid in a finite-dimensional Euclidean space. Since the objective function is continuous and the constraint defines a compact feasible set, the  Weierstrass Extreme Value Theorem (see, for example, Theorem 2.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}) ensures the existence of a global maximizer.

Define the Lagrangian:
\[
\mathcal{L}(\{\boldsymbol{\gamma}_j\}, \boldsymbol{\xi}_{K+1}, \mu) 
:= 
	\left( \sum_{j=1}^K \mathbf{u}_j^\top B \boldsymbol{\gamma}_j + \mathbf{v}^\top \boldsymbol{\xi}_{K+1} \right)^2
- 
\mu 
\biggl[
\bigl\{
	\sum_{j=1}^K \boldsymbol{\gamma}_j^\top (B + \lambda_j B^{\prime\prime}) \boldsymbol{\gamma}_j + 	\boldsymbol{\xi}_{K+1}^\top \boldsymbol{\xi}_{K+1} 
	\bigr\}
	- 1
	\biggr]
.
\]
The constraint function   is continuously differentiable, and its gradient vanishes only at the origin. Since the origin is not feasible, the gradient  is nonzero at all feasible points. Thus, the linear independence constraint qualification (LICQ) holds, ensuring that the KKT conditions are necessary for local maxima (see, for example, Theorem 5.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}).
Therefore, for any local maxima, gradients of $\mathcal{L}$ must be zero:
\begin{align*}
	\nabla_{\boldsymbol{\gamma}_j} \mathcal{L} &= 
	\frac{2}{n^2} 
	\left(
	\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \boldsymbol{\gamma}_\ell + \mathbf{y}^\top Z \boldsymbol{\xi}_{K+1}
	\right)
	B \Theta_j^\top \mathbf{y}
	- 2\lambda (B + \lambda_j B^{\prime \prime}) \boldsymbol{\gamma}_j = 0, \quad \forall j, 
	\numberthis \label{lag_grad_d}
	\\
	\nabla_{\boldsymbol{\xi}_{K+1}} \mathcal{L} &= 
	\frac{2}{n^2} 
	\left(
	\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \boldsymbol{\gamma}_\ell + \mathbf{y}^\top Z \boldsymbol{\xi}_{K+1}
	\right)
	Z^\top \mathbf{y}
	- 2\lambda \boldsymbol{\xi}_{K+1} = 0.
	\numberthis \label{lag_grad_zeta}
\end{align*}
By pre-multiplying equation \eqref{lag_grad_d} by $B^{-1}$,  and observing that  $2n^{-2}\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \boldsymbol{\gamma}_\ell + \mathbf{y}^\top Z \boldsymbol{\xi}_{K+1}$, which is a scalar quantity, appears in both \eqref{lag_grad_d} and \eqref{lag_grad_zeta}, the  conditions above together imply that any local maxima must be  scalar multiples of
$
\Theta_1^\top \mathbf{y}, \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y},
$
all scaled by the same factor. There exists only one such set of  vectors  that also satisfies the primal equality constraint, namely the normalized version of the vector set above. The global maximizer is therefore given by
\begin{equation*}
	\boldsymbol{\gamma}_j^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	\Theta_j^\top \mathbf{y},
	\quad
	%%%%%%%%%%%%%%%%
	\boldsymbol{\xi}_{K+1}^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	Z^\top \mathbf{y},
\end{equation*}

 	Observe that the objective is linear in each $\boldsymbol{\gamma}_j$ and in $\boldsymbol{\xi}_{K+1}$, and the constraint is a strictly convex quadratic form. Hence, the optimal solution lies in the direction of the gradient of the linear objective.

 	Define scalar weights $c_j$ and $c$ such that
 	\[
 	\boldsymbol{\gamma}_j = c_j \mathbf{u}_j, \quad \boldsymbol{\xi}_{K+1} = c \mathbf{v}.
 	\]

@@

 	\textbf{Conclusion}: The optimal coefficients are
 	\[
 	\boldsymbol{\gamma}_j = \frac{a_j}{d_j \sqrt{S}} \mathbf{u}_j, \quad \boldsymbol{\xi}_{K+1} = \frac{1}{\sqrt{S}} \mathbf{v},
 	\]
 	where
 	\[
 	a_j = \mathbf{u}_j^\top B \mathbf{u}_j, \quad d_j = \mathbf{u}_j^\top (B + \mu_j B^{\prime\prime}) \mathbf{u}_j, \quad \mathbf{u}_j = \Theta_j^\top \mathbf{y}, \quad \mathbf{v} = Z^\top \mathbf{y},
 	\]
 	and
 	\[
 	S = \sum_{j=1}^K \frac{a_j^2}{d_j} + \| \mathbf{v} \|^2.
 	\]

>>>>

	\subsection{Obtaining the Estimated Hybrid Regression Coefficient}
	\label{subsec: Obtaining the Estimated Hybrid Regression Coefficient}
	The regression coefficient $\boldsymbol{\eta}$ in model \eqref{eq: Hybrid functional model} can be written in terms of the estimated PLS components and scores. First, note that we can show $\widehat{\rho}_{il} = \langle \mathbf{W}_i^{[l]}, \widehat{\xi}_l \rangle_\mathcal{H} = \langle \mathbf{W}_i, \widehat{\boldsymbol{\zeta}}_l \rangle_\mathcal{H}$, where $\widehat{\boldsymbol{\zeta}}_l = \widehat{\xi}_l - \sum_{u=1}^{l-1} \langle \widehat{\boldsymbol{\delta}}_u, \widehat{\xi}_l \rangle_{\mathcal{H}} \widehat{\boldsymbol{\zeta}}_u$ for $l \ge 2$, with $\widehat{\boldsymbol{\zeta}}_1 = \widehat{\xi}_1$. Then, the decomposition of $Y_i$ in \eqref{eq: decomposition} leads to
	\begin{equation*}
		Y_i = \sum \limits_{l=1}^L \widehat{\nu}_l \langle \mathbf{W}_i^{[l]}, \widehat{\xi}_l \rangle_\mathcal{H} +\epsilon_i = \left\langle \mathbf{W}_i, \sum \limits_{l=1}^L\widehat{\nu}_l \widehat{\boldsymbol{\zeta}}_l \right\rangle_\mathcal{H}+\epsilon_i,
	\end{equation*}
	which, given the uniqueness of $\boldsymbol{\eta}$, leads to
	\begin{equation*}
		\widehat{\boldsymbol{\eta}} = \sum \limits_{l=1}^L\widehat{\nu}_l \widehat{\boldsymbol{\zeta}}_l.
	\end{equation*}
	Here, the number of PLS components $L$ to be estimated can be chosen by cross-validation.

@@

	Let $\xi = (\xi_1, \ldots, \xi_K, \boldsymbol{\zeta}) \in \mathcal{H}$. Assume that each functional component $\xi_j \in L^2([0,1])$ lies in the span of the basis functions ${b_1(t), \ldots, b_M(t)}$ and can be written as

@@

		\boldsymbol{\zeta} 

@@

		( \mathbf{Z}_i^\top \boldsymbol{\zeta} )

@@

		Z \boldsymbol{\zeta} 

@@

		\boldsymbol{\zeta}^\top \boldsymbol{\zeta}

@@

		\boldsymbol{\zeta}^\top \boldsymbol{\zeta},

@@

 \section{Proof of Proposition  \ref{proposition:linear_noregul}}\label{section:proof:proposition:linear_noregul}
 	The optimization problem
 	\ref{eq: Regularized generalized rayleigh quotient equation} can be written as
 		\max_{\boldsymbol{\gamma}_1, \ldots, \boldsymbol{\gamma}_K \in \mathbb{R}^M,
 			\boldsymbol{\zeta} \in \mathbb{R}^p
 		}~&
 		Z \boldsymbol{\zeta} 
 		\biggr)^2,
 		\\\text{subject to}\quad&	\sum \limits_{j=1}^K 
 	(	B + \lambda_k B^{\prime \prime})
 		\boldsymbol{\zeta}^\top \boldsymbol{\zeta} = 1.
 Let $\mathbf{u}_j := B \Theta_j^\top \mathbf{y} \in \mathbb{R}^M$ and $\mathbf{v} := Z^\top \mathbf{y} \in \mathbb{R}^p$. These are fixed problem data and not decision variables. Then the optimization problem becomes
 		\max_{\boldsymbol{\gamma}_1, \ldots, \boldsymbol{\gamma}_K \in \mathbb{R}^M,
 			\boldsymbol{\zeta} \in \mathbb{R}^p
 		}~&
 	\mathbf{u}_j^\top		
 		\mathbf{v}^\top \boldsymbol{\zeta} 
 		\biggr)^2,
 		\\\text{subject to}\quad&	\sum \limits_{j=1}^K 
 		(	B + \lambda_k B^{\prime \prime})
 		\boldsymbol{\zeta}^\top \boldsymbol{\zeta} = 1.
 	The objective function is continuous.
The feasible set is compact, as it corresponds to the boundary of an ellipsoid in a finite-dimensional Euclidean space. Therefore, the  Weierstrass Extreme Value Theorem (see, for example, Theorem 2.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}) ensures the existence of a global maximizer.
The constraint function is continuously differentiable, and its gradient vanishes only at the origin. Since the origin is not a feasible point, the gradient  is nonzero at all feasible points. Thus, the linear independence constraint qualification (LICQ) holds, ensuring that the KKT conditions are necessary for local maxima (see, for example, Theorem 5.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}).
To use this fact,
define the Lagrangian as follows:
\[
\mathcal{L}(\{\boldsymbol{\gamma}_j\}, \boldsymbol{\zeta}, \mu) 
:= 
	\left( \sum_{j=1}^K \mathbf{u}_j^\top  \boldsymbol{\gamma}_j + \mathbf{v}^\top \boldsymbol{\zeta} \right)^2
- 
\mu 
\biggl[
	\sum_{j=1}^K \boldsymbol{\gamma}_j^\top (B + \lambda_j B^{\prime\prime}) \boldsymbol{\gamma}_j + 	\boldsymbol{\zeta}^\top \boldsymbol{\zeta} 
	- 1
	\biggr]
.
\]
Then, if
$(\boldsymbol{\gamma}_1, \ldots, \boldsymbol{\gamma}_K, 
	\boldsymbol{\zeta}
)$
is a local maximizer, the gradients of $\mathcal{L}$ must vanish at that point:
	\nabla_{\boldsymbol{\gamma}_j} \mathcal{L} &= 
	s
	 \mathbf{u}_j
	- 2\mu (B + \lambda_j B^{\prime \prime}) \boldsymbol{\gamma}_j = 0, \quad \forall j, 
	\numberthis \label{lag_grad_d}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\nabla_{\boldsymbol{\zeta}} \mathcal{L} &= 
	s
\mathbf{v}
	- 2\mu \boldsymbol{\zeta} = 0,
	\numberthis \label{lag_grad_zeta}
where 
	s := \frac{2}{n^2} 
	\left(
	\sum_{\ell=1}^K \mathbf{u}_\ell^\top B \boldsymbol{\gamma}_\ell + \mathbf{v}^\top  \boldsymbol{\zeta}
	\right).
Since we assumed in Section~\ref{sec: finite basis approximation} that the functions $\{b_m(t)\}$ and their second derivatives are linearly independent, both $B$ and $B^{\prime\prime}$ are positive definite (see, for example, Theorem 273 of \citealp{gockenbachFiniteDimensionalLinearAlgebra2010}). As positive definiteness is preserved under conic combinations, $B + \lambda_j B^{\prime\prime}$ is also positive definite. Therefore, \eqref{lag_grad_d} implies that 
	=
	\frac{s}{2\mu}	
	(B + \lambda_j B^{\prime \prime})^{-1}
	\mathbf{u}_j
	   , \quad \forall j.
If $\boldsymbol{\gamma}_j$ is part of a local optimum, then it must be a scalar multiple of 
$\tilde{\mathbf{u}}_j := (B + \lambda_j B^{\prime \prime})^{-1} \mathbf{u}_j$, 
i.e., $\boldsymbol{\gamma}_j = c_j \tilde{\mathbf{u}}_j$.
Similarly, \eqref{lag_grad_zeta} implies that if 
$\boldsymbol{\zeta}$ is a local optimum, then it must be a scalar multiple of $\mathbf{v}$,
i.e., $\boldsymbol{\zeta} = d \mathbf{v}$.

Therefore, it suffices to 
By pre-multiplying equation \eqref{lag_grad_d} by $B^{-1}$,  and observing that  $2n^{-2}\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \boldsymbol{\gamma}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}$, which is a scalar quantity, appears in both \eqref{lag_grad_d} and \eqref{lag_grad_zeta}, the  conditions above together imply that any local maxima must be  scalar multiples of
$
\Theta_1^\top \mathbf{y}, \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y},
$
all scaled by the same factor. There exists only one such set of  vectors  that also satisfies the primal equality constraint, namely the normalized version of the vector set above. The global maximizer is therefore given by
\begin{equation*}
	\boldsymbol{\gamma}_j^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	\Theta_j^\top \mathbf{y},
	\quad
	%%%%%%%%%%%%%%%%
	\boldsymbol{\zeta}^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	Z^\top \mathbf{y},
\end{equation*}

 	Observe that the objective is linear in each $\boldsymbol{\gamma}_j$ and in $\boldsymbol{\zeta}$, and the constraint is a strictly convex quadratic form. Hence, the optimal solution lies in the direction of the gradient of the linear objective.

 	Define scalar weights $c_j$ and $c$ such that
 	\[
 	\boldsymbol{\gamma}_j = c_j \mathbf{u}_j, \quad \boldsymbol{\zeta} = c \mathbf{v}.
 	\]

@@

 	\textbf{Conclusion}: The optimal coefficients are
 	\[
 	\boldsymbol{\gamma}_j = \frac{a_j}{d_j \sqrt{S}} \mathbf{u}_j, \quad \boldsymbol{\zeta} = \frac{1}{\sqrt{S}} \mathbf{v},
 	\]
 	where
 	\[
 	a_j = \mathbf{u}_j^\top B \mathbf{u}_j, \quad d_j = \mathbf{u}_j^\top (B + \mu_j B^{\prime\prime}) \mathbf{u}_j, \quad \mathbf{u}_j = \Theta_j^\top \mathbf{y}, \quad \mathbf{v} = Z^\top \mathbf{y},
 	\]
 	and
 	\[
 	S = \sum_{j=1}^K \frac{a_j^2}{d_j} + \| \mathbf{v} \|^2.
 	\]
